---
title: "yuqi-FP-plot_trial"
author: "Yuqi Miao"
date: "11/12/2019"
output:
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r}
library(plotly)
library(shiny)
library(tidyverse)
library(rworldmap)
library(arsenal)
library(leaflet)
library(rworldmap) 
library(countrycode)  # Gets country code 
library(viridis)
```


# firstly draw every years map

```{r}
happy = read_csv("data/final_data_all_country.csv") %>% 
  janitor::clean_names() %>% 
  unique %>% 
  mutate(label = str_c("<b>Happiness: ", round(life_ladder,2), 
                      "</b><br>Country : ", country_name,
                      sep = ""),
        code = countrycode(country_name, 'country.name', 'iso3c'),## match code
        code = replace_na(code,"XKX"))

```

with long and lat, how to map in plotly?

1. not long and lat but world choroplath

plot the map according to year. Try to make it a function taking year as argument.
```{r}
# light grey boundaries
l <- list(color = toRGB("grey"), width = 0.5)

# specify map projection/options
g <- list(
  showframe = FALSE,
  showcoastlines = FALSE,
  projection = list(type = 'Mercator')
)

## making map according to year

global_happy_year <- function(i){
  happy_yeari = happy %>% 
    filter(year == i)
  p <- plot_geo(happy_yeari) %>%
  add_trace(
    z = ~life_ladder, color = ~life_ladder, colors = 'Blues',
    text = ~label, locations = ~code, marker = list(line = l)
  ) %>%
  colorbar(title = 'Happiness index') %>%
  layout(
    title = paste(as.character(i), "Global Happiness Index"),
    geo = g
  )
  return(p)
}

global_happy_year(2018)
global_happy_year(2017)
global_happy_year(2016)
global_happy_year(2011)



```
create a link?
# Create a shareable link to your chart
# Set up API credentials: https://plot.ly/r/getting-started
chart_link = api_create(p, filename="choropleth-world")
chart_link
2. with long and lat?
necessary?


# Secondly, try to add movable bar.

1. can we substitute this movie histogrom to our plot?
```{r}
ui <- shinyUI(fluidPage(
  titlePanel("Years"),
  sidebarPanel(
    sliderInput("year", "Year:", min = 2011, max = 2018, value = 1)
  ),
  mainPanel(
    plotlyOutput("trendPlot")
  )
))


## instead with world map 

server <- shinyServer(function(input, output) {

  output$trendPlot <- renderPlotly({

    ##plot function
    p <- global_happy_year(input$year)
    # style the xaxis
    layout(p)
  })
})
shinyApp(ui, server)
```

### yuqimiao add animation bar to the plot.

```{r}
Sys.setenv("plotly_username"="miaoyuqi")
Sys.setenv("plotly_api_key"="jqJ20PtvBRUpOBj2vlxb")
```


```{r}
# light grey boundaries
l <- list(color = toRGB("grey"), width = 0.5)

# specify map projection/options
g <- list(
  showframe = FALSE,
  showcoastlines = FALSE,
  projection = list(type = 'Mercator')
)
p <- plot_geo(happy) %>%
  add_trace(
    z = ~life_ladder, color = ~life_ladder, colors = 'Blues',
    text = ~label, locations = ~code, marker = list(line = l),frame = ~year
  ) %>%
  colorbar(title = 'Happiness index') %>%
  layout(
    title = "Global Happiness Index",
    geo = g
  )
p <- api_create(p, filename = "global_animation_plot")
p
```






















